Mixing time estimation in reversible Markov chains from a single sample path

### 2015-06-01 empirical.tex
minor tweak to discussion of group inverse

We adopt a different strategy for estimating $\vpi$, which leads to
our construction of empirical confidence intervals, detailed in
\cref{alg:empest}.
We form the matrix $\wh\vP$ using smoothed frequency estimates of
$\vP$ (Step~\ref{step:P}), then 
compute the so-called group inverse $\giAh$ of $\wh\vA = I - \wh\vP$,
followed by finding the unique stationary
distribution $\hat\vpi$ of $\wh\vP$ (Step~\ref{step:pi}).
The group inverse $\giAh$ of $\wh\vA$ is uniquely defined and if $\wh\vP$ defines an ergodic chain
(which is the case here due to the use of the smoothed estimates),
$\giAh$ can be computed at the cost of inverting an $(d-1)\times (d-1)$ matrix
(cf. Theorem 5.2 in \cite{meyer1975role}).
Further, once given $\giAh$, the unique stationary distribution of $\wh\vP$ can be 
read out from the last row of $\giAh$ (Theorem 5.3 of  \cite{meyer1975role}).
The group inverse will also be used later to compute the sensitivity of $\vpi$
based on a result from \cite{cho2001comparison}.
Based on $\hat\vpi$ and $\wh\vP$, we construct the plug-in estimate
$\wh\vL$ of $\vL$, and use the eigenvalues of its symmetrization to
form the estimate $\hatgap$ of the spectral gap (Steps~\ref{step:eig}
and~\ref{step:gap}).
In the remaining steps, 
we use perturbation analyses to relate $\hat\vpi$ and $\vpi$, viewing
$\vP$ as the perturbation of $\wh\vP$; and also to relate $\hatgap$
and $\gap$, viewing $\vL$ as a perturbation of $\Sym(\wh\vL)$.
Both analyses give error bounds entirely in terms of observable
quantities, tracing back to empirical error bounds for the smoothed
frequency estimates of $\vP$.
Note that the most expensive step in the algorithm is the computation of the group inverse,
which, as noted reduces to matrix inversion.
Thus, with a standard implementation of matrix inversion, the algorithm's complexity becomes 
$O( n + d^3)$, while the space complexity is $O(d^2)$.
\todoc{Implement and test?}

>>>>

We adopt a different strategy for estimating $\vpi$, which leads to
our construction of empirical confidence intervals, detailed in
\cref{alg:empest}.
We form the matrix $\wh\vP$ using smoothed frequency estimates of
$\vP$ (Step~\ref{step:P}), then compute the so-called group inverse
$\giAh$ of $\wh\vA = I - \wh\vP$ (Step~\ref{step:gi}), followed by
finding the unique stationary distribution $\hat\vpi$ of $\wh\vP$
(Step~\ref{step:pi}).
The group inverse $\giAh$ of $\wh\vA$ is uniquely defined; and if
$\wh\vP$ defines an ergodic chain (which is the case here due to the
use of the smoothed estimates), $\giAh$ can be computed at the cost of
inverting an $(d-1)\times (d-1)$ matrix~\citep[Theorem
5.2]{meyer1975role}.
Further, once given $\giAh$, the unique stationary distribution
$\hat\vpi$ of $\wh\vP$ can be read out from the last row of
$\giAh$~\citep[Theorem 5.3]{meyer1975role}.
The group inverse will also be used later to compute the sensitivity
of $\vpi$ based on a result from \cite{cho2001comparison}.
Based on $\hat\vpi$ and $\wh\vP$, we construct the plug-in estimate
$\wh\vL$ of $\vL$, and use the eigenvalues of its symmetrization to
form the estimate $\hatgap$ of the spectral gap (Steps~\ref{step:eig}
and~\ref{step:gap}).
In the remaining steps, 
we use perturbation analyses to relate $\hat\vpi$ and $\vpi$, viewing
$\vP$ as the perturbation of $\wh\vP$; and also to relate $\hatgap$
and $\gap$, viewing $\vL$ as a perturbation of $\Sym(\wh\vL)$.
Both analyses give error bounds entirely in terms of observable
quantities, tracing back to empirical error bounds for the smoothed
frequency estimates of $\vP$.
Note that the most expensive step in the algorithm is the computation of the group inverse,
which, as noted reduces to matrix inversion.
Thus, with a standard implementation of matrix inversion, the algorithm's complexity becomes 
$O( n + d^3)$, while the space complexity is $O(d^2)$.
\todoc{Implement and test?}
